### 2020-01-26 introduction.tex

It is the combination of differential comparison and mutation that is key.  Differential comparison of tools, as noted above, is not really meaningful, without additional effort; na\"ive methods simply will not work \cite{regehrRandom}.  Consider a comparison of the number of findings between two tools over a single program, or over a large set of programs.  If one tool emits more warnings and errors than another, it may mean that the tool is more effective at finding bugs; but it may also mean that it has a higher false positive rate.  Without human examination of the individual findings, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.  Using mutants, however, provides a foreground to compare to this background.  In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer findings on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.  This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.  While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated contracts while still detecting \emph{more} mutants.  Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?

>>>>

It is the combination of differential comparison and mutation that is key.  Differential comparison of tools, as noted above, is not really meaningful, without additional effort; na\"ive methods simply will not work \cite{regehrRandom}.  Consider a comparison of the number of findings between two tools over a single program, or over a large set of programs.  If one tool emits more warnings and errors than another, it may mean that the tool is more effective at finding bugs; but it may also mean that it has a higher false positive rate.  Without human examination of the individual findings, it is impossible to be sure, or even (in cases where the tools are reasonably comparable) to make an informed guess.  Using mutants, however, provides a foreground to compare to this background.  In particular, for a large set of programs, the most informative result will be when 1) tool A reports fewer findings on average than tool B over the un-mutated programs but 2) tool A also detects more mutants.  This is strong evidence that A is simply better all-around than B; it likely has a lower false positive rate \emph{and} a lower false negative rate.  While it is not proof of this claim, it is hard to construct another plausible explanation for reporting \emph{fewer} findings on un-mutated code while still detecting \emph{more} mutants.  Other than having better precision and recall, how else could a tool effectively distinguish mutated from un-mutated code?

### 2020-01-26 relatedwork.tex

Klinger et al. propose a different approach to differential testing of analysis tools \cite{klinger2019differentially}.  Their approach is in some ways similar to ours, in that it takes as input a set of seed programs, and compares results across new versions generated from that seed.  The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available) and their tool must parse and understand the programs, and that the new versions are based on adding new assertions, not ``breaking'' the original code.  We allow arbitrarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.  Further, their approach only identifies problems when tools are outliers compared to numerous other tools in either detecting a bug (precision) or not detecting it (soundness), and so requires comparing multiple tools.  Our approach has some utility for even a single tool (you can just examine prioritized un-detected mutants).  On the other hand, their approach can identify precision issues, while we offer no real help with false positives (in theory, you could apply their majority-vote method to mutants only a few tools flag, but mutants \emph{are} usually faults. in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).  Most importantly, however, their approach only applies to tools that check assertions, rather than more general static analysis tools that may only identify bad code patterns, not compute state reachability (e.g, most of the widely used tools).

Finally, the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.  Methods for improving mutation analysis in testing will often (but not always, e.g. those based on coverage or other dynamic information) transfer to our context as well.

>>>>

Klinger et al. propose a different approach to differential testing of analysis tools \cite{klinger2019differentially}.  Their approach is in some ways similar to ours, in that it takes as input a set of seed programs, and compares results across new versions generated from that seed.  The primary differences are that their seed programs must be warning-free (which greatly limits the set of input programs available) and their tool must parse and understand the programs, and that the new versions are based on adding new assertions, not ``breaking'' the original code.  We allow arbitrarily buggy seed programs (thus many more real programs can be used), and can, due to the any-language nature of the mutation generator we use, operate even in new languages without further development effort.  Further, their approach only identifies problems when tools are outliers compared to numerous other tools in either detecting a bug (precision) or not detecting it (soundness), and so requires comparing multiple tools.  Our approach has some utility for even a single tool (you can just examine prioritized un-detected mutants).  On the other hand, their approach can identify precision issues, while we offer no real help with false positives (in theory, you could apply their majority-vote method to mutants only a few tools flag, but mutants \emph{are} usually faults. in contrast to their introduction of checks that may be guaranteed to pass, so this is probably not very helpful).  Most importantly, however, their approach only applies to tools that check assertions, rather than more general, popular static analysis tools that only identify bad code patterns.

Finally, the large body of work on using mutants in software testing \cite{jia2011analysis,demillo1978hints,budd1980theoretical, groce2015verified,groce2018verified,MutGoogle,ivankovic2018industrial,mutKernel} is obviously relevant, in that we basically adopt its approach, but re-define killing a mutation for a static analysis context.
